---
title: "purrr_and_dataframes"
author: "Jackie Lu"
date: "10/31/2020"
output:
  html_document:
    css: "dark_mode.css"
---

# Purrr and dataframes

|     Applying functions to elements is a pretty common situation in programming. In R, the purrr library allows you to do iterations with consistent syntax plus <a href="https://stackoverflow.com/a/47123420">other benefits</a>.  

<br>

|     While using purrr though, I found it difficult to remember which function was used for columns of a data frame, and which was used for rows of a data frame. I decided to investigate.

<br>


```{r}
# # Use code below if you haven't installed purrr yet
# install.packages("purrr")

library(purrr)

# Create a data frame of values
vals_1 <- data.frame(a_col = c(1, 2, 3),
                     b_col = c(1, 2, 3))
```

You may have noticed at some point that you can access the elements of a dataframe by using double square brackets like so:
```{r}
# Access the first element in the vals_1 list
vals_1[[1]]
```

Data frames are actually a type of list where the elements are vectors, and the lengths of these vectors are the same. The data type within a single vector is consistent as well. Using:

```{r}
as.list(vals_1)
```

We can see that a data frame is a list of its columns. In the purrr cheatsheet
it mentions that $\texttt{map}$ "[applies] a function to each element of a list." Clearly, $\texttt{map}$ must be used for columns then, as the data frame is the list and the columns are its elements.

```{r}
# Go through each column in vals_1, and sum the column
map(vals_1, sum)
```

The description for pmap says that it "[applies] functions to groups of elements from list of lists, vectors." If we look at the example below, we see that every element in a column corresponds to a row in that column.  

```{r}
# First element in the a column
vals_1$a_col[1]
# Second element in the a column
vals_1$a_col[2]
# Third element in the a column
vals_1$a_col[3]
```

If we have the first element of every column, then we have the first row.

```{r}
# Vector of first elements
c(vals_1$a_col[1], vals_1$b_col[1])

# First row of the data frame of values
vals_1[1, ]
```

We see that we get the same values. We can think of rows as groups of elements from a vector within a list. Therefore $\texttt{pmap}$ applies to rows of a dataframe.  

```{r}
pmap(vals_1, sum)
```

$\texttt{modify_if}$ is similar to $\texttt{map}$ in that it works on elements of a list, except it returns the same data type that was given to it, and only acts on the elements that fit the condition. In the code below, I show an example of using $\texttt{modify_if}$ for cleaning "?" into NA values.

```{r}
# Small example dataframe
ex1 <- data.frame(id = c(1, 2, 3),
                  value = c(NaN, 2, NA),
                  name = c("a", "b", "?"),
                  name_2 = c("?", "?", "c"))

replace_q_mark <- function(fact_vect, target_char = "?"){
  # Function that takes a factor vector and a character variable such as
  # "not applicable" and changes the factor levels that correspond to the
  # character variable so that it represents NA
  
  # Check if char_vect has a question mark
  if (sum(levels(fact_vect) == target_char) >= 1){
    # Replace it with NA
    levels(fact_vect)[levels(fact_vect) == target_char] <- NA
  }
  return(fact_vect)
}

# This returns a dataframe where the question marks are replaced with NA's
modify_if(ex1, is.factor, ~ replace_q_mark(.x, target_char = "?"))
```



## Resources

<section>
  <a href="https://rstudio.com/resources/cheatsheets/">
    R cheatsheets
  </a>
  <p>
    You can find the cheatsheet for purrr in here, among other R cheatsheets.
  </p>
</section>

<section>
  <a href = "https://jql6.github.io" class="end">
    Return to homepage
  </a>
</section>

<br>